---
title: "Assignment-B1"
author: "Ian Coccimiglio"
date: "2023-11-03"
output: github_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(testthat)
```

## Creating a non-zero counter function
The function countNonZero has a single input of a generic vector, and returns the number of non-zero entries. Zeros here are loosely defined, and include anything that is 'falsy' - such as 0, "0", and FALSE. The function stops if any entries are equal to NA. It runs using only base R, aside from function testing.

```{r}
#' Non-zero element counter
#' This function returns the number of non-zero elements in a vector 
#' (including values that can be coerced to equal 0, such as FALSE and "0").
#' It is also capable of handling non-vector inputs, such as lists and dataframes. 
#' Values of NA will throw an error as they cannot be assigned a value of true/false.
#' 
#' @param vec A vector of inputs. 
#' The name suggests that this function is designed to operate on generic vectors, 
#' and not explicitly designed for higher level structures.
#'
#' @return numNonZero An integer count of non zero values
#' @export
#'
#' @examples countNonZero(seq(0,5))
countNonZero <- function(vec) {
  if (is.data.frame(vec)) {
    print("Note: Counting non-zero entries across all rows and columns")
  }
  if (any(is.na(vec))) stop("NAs detected in input")
  nonZeroElements <- (vec != 0)
  numNonZero <- sum(nonZeroElements)
  return(numNonZero)
}

```

## Examples
This function is very flexible and works on a variety of inputs. In its simplest usage, this function takes in an ordinary numeric vector, and returns the number of non-zero elements. 

```{r}
ordinary_vector <- c(0, 0, 5, 2, 3, 0, 1)
countNonZero(ordinary_vector)
```

However, the function handles non-typical inputs as well, such as string vectors and boolean vectors:

```{r}
string_vec <- countNonZero(c("0", "0", "1", "2"))
bool_vec <- countNonZero(c(FALSE, TRUE, TRUE, TRUE, FALSE))
print(paste("String Vector =",string_vec))
print(paste("Boolean Vector =",bool_vec))
```

In typical usage, you might have a problem like "counting the number of students who have first aid training", where there is a binary column in a dataframe. We can easily use countNonZero to assess the number of students who have this training.

```{r}
df <- data.frame('Name'=c("John", "Sarah", "Jeff", "Karen", "James"), "First_Aid_Training"=c(0,1,1,0,1))
countNonZero(df$First_Aid_Training)
```

In more niche use cases, you can count the total number of zeros present within a dataframe across all rows and columns. This prints a note so the user knows what's happening.

```{r}
df <- data.frame('Name'=c("John", "Sarah", "Jeff", "Karen", "0"), "First_Aid_Training"=c(0,1,1,0,1))
countNonZero(df)
```

## Tests

```{r}
df <- data.frame('Name'=c("John", "Sarah", "Jeff", "Karen", "Jeff"), "First_Aid_Training"=c(0,1,1,0,1))

test_that("numeric vectors work", {
  expect_equal(countNonZero(c(0, 0, 0)), 0)
  expect_equal(countNonZero(c(1, 1, 1)), 3)
  expect_equal(countNonZero(c(0, 2, 2)), 2)
})

test_that("string vectors work", {
  expect_equal(countNonZero(c("0", "0", "0")), 0)
  expect_equal(countNonZero(c("0", "1", "1")), 2)
})

test_that("NAs create errors", {
  expect_error(countNonZero(c(NA, NA)))
  expect_error(countNonZero(c(1, 0, NA)))
})

test_that("Dataframes are counted across all columns", {
  expect_equal(countNonZero(df), 8)
})

```